--- 
title: "16s rRNA analysis workshop"
author: "Hena R. Ramay"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This tutorial covers workshop material for all three days"
---

# Introduction

Placeholder


## Workshop Schedule
## Important links

<!--chapter:end:index.Rmd-->


# Day One

Placeholder


## What are we trying to achieve
## Basics & Background
## DADA2 pipeline (v1.2)
### Data for the tutorial
### Getting ready ( load packages and get file list)
### Inspect read quality profiles
### 

<!--chapter:end:02-Day1.Rmd-->

# Day Two

## Main concepts to be discussed:
- Finish dada2 pipeline
- Assign Taxonomy
- Intro to Phyolseq package
- Create a Phylum level bar plots
- Alpha diversity plots
- Beta diversity plots


Once we are finished using the dada2 package, we will have a sequence table and taxonomy table. 

Lets look at the metadata files we have to get more information about these samples.

```{r,metadata}

library(ggplot2)
library(dplyr)
time_info<-read.delim("~/projects/16sanalysis_workshop/workshoptutorial/MiSeq_SOP/mouse.time.design",sep  = "\t")
dpw_info<-read.delim("~/projects/16sanalysis_workshop/workshoptutorial/MiSeq_SOP/mouse.dpw.metadata",sep  = "\t")

sample_info<-left_join(time_info,dpw_info,by="group")

rownames(sample_info)<- sample_info$group

sample_info
```

Lets make a phyloseq object

```{r}




```

Now that we have sample_info lets try to make a phyloseq object out of this

```{r, phyloseq}
library(phyloseq)
taxa<-readRDS("~/projects/16sanalysis_workshop/workshoptutorial/output/taxa.rds")
seq<-readRDS("~/projects/16sanalysis_workshop/workshoptutorial/output/seq.rds")

ps <- phyloseq(otu_table(seq,taxa_are_rows = F), 
               sample_data(sample_info), 
               tax_table(taxa))

ps
```

You can see that this object has an OTU table(ASV table), sample data and tax_table. You can use functions tax_table(), sample_data() and otu_table() to access the data.

Take a look at:

- subset_samples()
- subset_taxa()
- tax_glom()
- sample_sums()
- prune_samples()
- transform_sample_counts()
- psmelt()



```{r}
ps2<-tax_glom(ps,taxrank = "Phylum")
ps2 = transform_sample_counts(ps2, function(x) x/sum(x))
pmelt<-psmelt(ps2) %>% arrange(desc(Abundance))

cutoff<-0.005
pmelt_filt<-pmelt %>% group_by(Phylum,time) %>% filter(sum(Abundance) >cutoff)

ggplot(pmelt_filt,aes(x=Sample,y=Abundance,color=Phylum,fill=Phylum)) +geom_bar(stat = "identity")

pmelt_filt %>% group_by(time,Phylum) %>% summarise(mean_Abundance=mean(Abundance)) %>% ggplot(.,aes(x=time,y=mean_Abundance,fill=Phylum)) +geom_bar(stat="identity")

```

Figure out how to change the colors to custom colors!!

```{r, summary}


pmelt_filt %>% group_by(time,Phylum) %>% summarise(mean=mean(Abundance))

```


## Alpha & beta Diversity




```{r, alpha diversit,echo=F}

richness<-estimate_richness(ps)

ggplot(richness,aes(x=group,y=Shannon,color=group,fill=group)) + geom_boxplot()

```

```{r,echo=F}
ps %>%
transform_sample_counts(., function(x) log(1 + x)) %>%
plot_ordination(., ordinate(., method="PCoA", distance="bray"),
color="group" ,title="Bray NMDS")+ scale_colour_manual(values=cols) + geom_point()#%>% #+ theme_bw() 

```


<!--chapter:end:03-Day2.Rmd-->

# Day Three

<!--chapter:end:04-Day3.Rmd-->

# Summary


<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

